A Dependently Typed Multi-Stage Calculus

### 2019-08-30 text/introduction.tex
For arxiv.

\subsection{Our Work}
In this paper, we develop a new multi-stage calculus \LMD by extending
the existing multi-stage calculus \LTP\cite{Hanada2014} with dependent
types and study its properties.  We base our work on \LTP, in which
the four multi-stage constructs are handled slightly differently from
MetaOCaml, because its type system and semantics are arguably simpler
than \(\lambda^\alpha\)~\cite{taha2003environment}, which formalizes
the design of MetaOCaml more faithfully.  Dependent types are based on
\LLF~\cite{attapl}, which has one of the simplest forms of dependent
types.  Our technical contributions are summarized as follows:
\begin{itemize}
\item We give a formal definition of \LMD with its syntax, type system and
two kinds of reduction: full reduction, allowing reduction of any redex,
including one under $\lambda$-abstraction and quotation, and staged reduction, a
small-step call-by-value operational semantics that is closer to the intended
multi-stage implementation.
\item We show preservation, strong normalization, and confluence for
  full reduction; and show unique decomposition (and progress as its
  corollary) for staged reduction, which formalizes program execution.
\end{itemize}
The combination of multi-stage programming and dependent types has
been discussed by Pasalic, Taha, and Sheard~\cite{pasalic2002tagless}
and Brady and Hammond~\cite{brady2006dependently} but, to our
knowledge, our work is a first formal calculus of dependently typed
multi-stage programming with all the key constructs~\cite{MetaML}.

>>>>

\subsection{Our Work}
In this paper, we develop a new multi-stage calculus \LMD by extending
the existing multi-stage calculus \LTP\cite{Hanada2014} with dependent
types and study its properties.  We base our work on \LTP, in which
the four multi-stage constructs are handled slightly differently from
MetaOCaml, because its type system and semantics are arguably simpler
than \(\lambda^\alpha\)~\cite{taha2003environment}, which formalizes
the design of MetaOCaml more faithfully.  Dependent types are based on
\LLF~\cite{attapl}, which has one of the simplest forms of dependent
types.  Our technical contributions are summarized as follows:
\begin{itemize}
\item We give a formal definition of \LMD with its syntax, type system and
two kinds of reduction: full reduction, allowing reduction of any redex,
including one under $\lambda$-abstraction and quotation, and staged reduction, a
small-step call-by-value operational semantics that is closer to the intended
multi-stage implementation.
\item We show preservation, strong normalization, and confluence for
  full reduction; and show unique decomposition (and progress as its
  corollary) for staged reduction.
\end{itemize}
The combination of multi-stage programming and dependent types has
been discussed by Pasalic, Taha, and Sheard~\cite{pasalic2002tagless}
and Brady and Hammond~\cite{brady2006dependently} but, to our
knowledge, our work is a first formal calculus of full-spectrum dependently typed
multi-stage programming with all the key constructs mentioned above.
